\chapter{Acknowledgements}

This manuscript and my journey as a PhD student wouldn't have been the same without the steady presence and contribution of many wonderful people.

\bigskip

I'm extremely grateful to the members of my jury --- Slavica Jonic, Charles Kervrann, Gabor Papai, Leandro Estrozi, and Marc Jamin --- for offering their time and expertise in reading this manuscript and for the following discussion at the thesis defense.

Many thanks to GRAL for providing the funding and resources that gave me a smooth PhD experience. In particular, heartfelt thanks to Anne-Mathilde Thierry for her rare ability to make interactions with the cumbersome french bureaucracy quick and pleasant.

\bigskip

There are many people I have worked with and learned from, at IBS and outside, that have helped me during these years.

My most heartfelt thanks go to my supervisors Irina and Joanna. I'm truly grateful for your continued support towards my independence and self-determination, all while managing to direct my (admittedly often unfocused) efforts towards useful goals. I was incredibly lucky to work with scientists who so value collaboration, curiosity, open discussion, and good scientific and work ethic. You created a stimulating and friendly environment that I will forever crave going forward.

I'm deeply indebted to Alister, for his inexhaustible willingness to share knowledge, for the long discussions on software design, and for giving me a head-start in the world of Cryo-ET. Thank you also for introducing me to the napari community, which so much shaped the next few years of both of our careers. 

Special thanks to Leandro, for being always ready to explain and re-explain complex image processing topics, and for easily sharing his no-nonsense yet lighthearted attitude towards pretty much anything.

Thanks Ambroise, for all the technical discussions on cryo-EM processing and its nuances, especially regarding helical assemblies, and for the occasional evening party.

Many thanks also to Maria, Madalen, Irina and Joanna for taking care of the experimental side of this thesis; I couldn't have done this without you!

I also had great pleasure working with Linda Sandblad and everyone else at the Centre for Electron Microscopy in Umeå. I had a great time there, doing both lots of science and sport with such a friendly group!

I'm extremely grateful to everyone in the napari community --- particularly the core developers, with whom I worked closely by now for several years. You've been an endless source of new ideas and of personal growth both as a developer and as a scientist. I look forward to continue working with you.

\bigskip

My deepest thanks to everyone who loves me and shared these last few years with me.

A big thanks to my parents Tiziana and Giancarlo and my sister Valentina, who always encouraged my studies unconditionally --- even when I suddenly chose to move abroad and through any financial challenges --- and who remain unwaveringly supportive of who I am and of everything I do.

Thank to Moritz for always joining in and enabling my unnecessary pedantry, and for readily accepting to be my first test subject whenever I developed something new.

Thank you Alessio for always preventing me from indulging in too much unnecessary pedantry, and for readily accepting to be my first test subject whenever I developed something new.

Special thanks Eymeline, for all that we shared at work and outside, and for showing me a plethora of new shiny nerdy things I didn't know about. Thank you for being so stubbornly selfless with everyone around you.

Thanks Paco for being my trusty board games companion, even when you're exhausted from work. Thanks for being always openhearted, and for unfailingly joining me in the fight against the System. 

Thanks Lorene, for all the philosophical discussions and for the good times in Grenoble and Umeå. I'm constantly inspired by your hard work and determination, all while maintaining a relaxed attitude that makes it easy to be around you.

Thank you also Jahnavi, for the always thoughtful gifts and for being a steady, quiet and friendly presence in the office. Good luck with finishing up!

Thank you Maaike, for always showing me by example how to have a more positive impact on the world and the people around me, and for navigating uncharted territories with me.

Thanks to all my friends back in Rovereto, for providing a safe haven at the Stube whenever I needed to take a break and unwind while playing games and eating pizza.

\bigskip

Thank you all!
